---
title: "ASSIGNMENT 5"
author: "Joshua Greenert"
date: '7/23/2022'
output:
  word_document: default
  html_document: default
  pdf_document: default
---
# Set the working directory.
```{r setup}
knitr::opts_knit$set(root.dir = normalizePath("C:/Users/Josh/Documents/GitHub/R-Projects/Statistics For Data Science"))
```

```{r include=FALSE, echo=FALSE}
# libraries
library(rmarkdown)
library(knitr)
library(ggplot2)
library(ggm)
theme_set(theme_minimal())
```
# Show the head of the data.
```{r echo=FALSE}
surveyResults = read.csv("data/student-survey.csv")
head(surveyResults)

# Show the covariance.


```
# Show the covariance
```{r echo=FALSE}
covarianceAll = cov(surveyResults)
covarianceAll

```
Using this calculation allows us to see the relationships between the data.  When there is a positive
covariance, the two sets of data appear to move together; whereas, if the covariance is negative,
the two sets of data move away from one another.  The results indicate that the variables TimeTV and 
Happiness appear to have a positive covariance while TimeReading and Happiness have a negative covariance.

# Examine the Survey data variables
**TimeReading** - appears to be in units of hours per week.\
**TimeTV** - appears to be in units of hours per week.\
**Happiness** - appears to be a percentage of overall happiness reported.\
**Gender** - appears to be binary between male or female.\

Since both measurements of time are in the same units, the covariance would not be impacted on the result
of comparing the two measurements.  When comparing these against a percentage, however, there may be a
significant difference in the output expected.  Nevertheless, the variables are suitable as they are since
they show a decent reflection of the dataset.

# Choose a correlation test to perform: pearson
This test is a wise choice as it measures the strength of the linear relationship between normally
distributed variables and is the default for most correlation tests.

# Perform a correlation analysis
## All Variables
```{r echo=FALSE}
correlationAll = cor(surveyResults)
correlationAll
```
## Single correlation between pair of variables
```{r echo=FALSE}
tvHappyCor = cor(surveyResults$TimeTV, surveyResults$Happiness)
readingHappyCor = cor(surveyResults$TimeReading, surveyResults$Happiness)
print('Correlation between tv and happiness')
tvHappyCor
print('Correlation between reading and happiness')
readingHappyCor
```
## Repeat the correlation test with confidence internal at 99%
```{r echo=FALSE}
print('Correlation between tv and happiness with Confidence intervals')
tvHappyCorTest = cor.test(surveyResults$TimeTV, surveyResults$Happiness, conf.level = 0.99)
tvHappyCorTest

print('Correlation between reading and happiness with Confidence intervals')
readingHappyCorTest = cor.test(surveyResults$TimeReading, surveyResults$Happiness, alternative = "less", conf.level = 0.99)
readingHappyCorTest
```
## Describe what the calculations suggest
The correlation tests provided data that suggested it was likely that there was both a negative and positive 
correlation among the variables chosen.  Further tests with confidence intervals also suggest that there is a 
positive correlation between tv time and happiness, while there is a negative correlation between reading and 
happiness.  

# Calculate the correlation coefficient and coefficient of determination; describe the results
```{r echo=FALSE}
print('Correlation coefficient between tv and happiness')
(tvHappyCor^2) * 100
print('Correlation coefficient between reading and happiness')
(readingHappyCor^2) * 100
```
From the data above, we can assess that tv is positively correlated with happiness but that it can 
only account for 40.52% of the variation in terms of happiness.  Alternatively, we can assess that 
reading is negatively correlated with happiness and that it can only account for 18.91% of the variation
in terms of happiness.  There remains 59.48% variability for tv and 81.09% variability that is unexplained
by the two samples provided.

# Can you say that watching more TV caused students to read less?
The data that is provided doesn't conclude whether Tv engagement effected reading.  Instead, the data
suggests happiness may not be as prevalent within reading as it is in TV.

## Partial Correlation Test with "reading" set as the control
```{r echo=FALSE}
partialCorrTest = pcor(c("TimeTV", "Happiness", "TimeReading"), var(surveyResults))
partialCorrTest
print('The result squared')
partialCorrTest^2
```
From the data, we can determine that reading accounted for approximately 35.72% of the variability 
between happiness and TV.  Therefore, it does have an effect on the results and suggests that TV usage 
may cause students to read less.  












